---
title: "onsetsync demo"
author: "Tuomas Eerola"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Onsetsync demo as Rmd notebook**

A Rmarkdown notebook with code examples in R for [onsetsync](https://github.com/tuomaseerola/onsetsync) by [Tuomas Eerola](https://www.durham.ac.uk/staff/tuomas-eerola/), [Music and Science Lab]() at [Durham University](https://www.durham.ac.uk) and [Music and Science Lab](https://musicscience.net).

To run the code in your RStudio, download the Rmd file and open it in your Studio. Here's the link: [https://github.com/tuomaseerola/onsetsync/onsetsync.Rmd](https://github.com/tuomaseerola/onsetsync/onsetsync.Rmd). It assumes that you have `devtools` package installed to be able to install packages from github.

File `onsetsync.Rmd` | Version `9/11/2022` | [Back to Index](https://github.com/tuomaseerola/onsetsync/)

# Load libraries

```{r}
library(ggplot2,quietly = TRUE)
library(tidyverse,quietly = TRUE)
require(devtools)
devtools::install_github("tuomaseerola/onsetsync")
library(onsetsync,quietly = TRUE)
ver <- packageVersion('onsetsync')
print(ver)
```

## Load data (internal)

Here we take a Cuban Son performance titled *Palo Santo* from *IEMP* dataset [@Clayton2020emr] at [https://osf.io/sfxa2/](https://osf.io/sfxa2/):

```{r}
CSS_Song2 <- onsetsync::CSS_IEMP[[2]]   # Read one song from internal data
CSS_Song2 <- dplyr::select(CSS_Song2,Label.SD,SD,Clave,Bass,Guitar,Tres,
                           Cycle,Isochronous.SD.Time) # Select some columns
print(head(CSS_Song2))
```

## Load Data -- From OSF (optional)

```{r}
require(httr)                           # Additional library needed
CSS_Song2_Onset <- get_OSF_csv('8a347') # Get onsets
print(head(CSS_Song2_Onset[,1:8,]))
```

## Summarise Raw Data

```{r}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres'))
print(knitr::kable(tab1,format = "markdown", digits = 1,
     caption = 'Onsets counts (N) and mean IOIs in ms.'))
```

## Visualisation

## Visualise Onset Structures

```{r}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual = 'Isochronous.SD.Time',
                     pcols = 2)
print(fig1)
```

## Visualise Deviations from Isochrony

```{r}
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Bass','Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE)
print(fig2)
```

## Visualise Across Sub-beats

```{r}
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE, box = TRUE)
print(fig2)
```

# Synchrony

## Classic Measures of Synchrony

See definitions in classic literature (Rasch, Wing, Clayton).

```{r}
d <- sync_sample_paired(CSS_Song2,'Clave','Bass',N=0,beat = 'SD')
s <- summarise_sync(d)
S <- data.frame(s); St<-t(S); colnames(St) <- 'ms'
knitr::kable(St,digits = 1,caption = 'Classic asynchrony measures.')
```

## Synchrony Between Instruments

Number of _shared onsets_ may vary across instruments/beats. Comparisons with `sampling` and `bootstrapping`.

```{r}
set.seed(1201) # set random seed
N <- 100       # Let's select random 100 onsets from both
Boot <- 1      # No multiple bootstraps
d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass', N, Boot,'SD',TRUE)
print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
```

## Synchrony Between Instruments (plot)

```{r}
inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
print(plot_by_pair(dn))  # plot
```

## Descriptive Statistics

```{r}
table3 <- data.frame(summarise_sync_by_pair(dn))
print(knitr::kable(table3,
                   digits = 1,
                   caption = 'Descriptives, t-test and p-value.'))
```


## Synchrony with other Variables

Is there a link between beat durations and synchrony?

```{r}
CSS_Song2 <- CSS_Song2 %>% 
  group_by(Cycle) %>%
  mutate(Duration = max(Isochronous.SD.Time) - min(Isochronous.SD.Time))
CSS_Song2 <- ungroup(CSS_Song2) # drop grouping structure
d1 <- sync_sample_paired(df = CSS_Song2,
                         INSTR1 = 'Clave', INSTR2 = 'Bass',
                         N = 200, BNum = 1, beat = 'Duration')
fig6 <- plot_by_variable(d1,meta = 'Clave-Bass Synchrony',xlab = 'Cycle duration (in s)')
print(fig6)
```

## Synchrony with other Variables

Are there specific trends in synchrony across the performance?

```{r}
d2 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Duration')
d3 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Isochronous.SD.Time')
tmp <- data.frame(asynch=d2$asynch*1000,Duration=d2$beatL,Time=d3$beatL)
fig9 <- plot_by_var_time(df = tmp,
                          var1 = 'Time',
                          var2 = 'asynch',
                          var3 = 'Duration',
                          ylabel = 'Asynchrony (ms)')
fig9
```


## Synchrony across Several Performances (corpus)

```{r}
corpus <- onsetsync::CSS_IEMP                              # This is a small corpus
D <- sync_sample_paired(corpus,'Tres','Bass',N=0,beat='SD')
D <- D$asynch
D$asynch_abs <- abs(D$asynch)*1000                         # Convert to ms
fig7 <- plot_by_dataset(D,'asynch_abs','name', box = TRUE)
print(fig7)
```

# Periodicity

## Periodicity of Guitar Onsets

```{r}
extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
fig10 <- gaussify_onsets(extract$Guitar, wlen = 0.2, plot = TRUE)
print(fig10$fig)
```

## Estimate Periodicity

Four methods to estimate periodicity (autocorrelation, FFT, smoothed FFT, or difference)

```{r, warning=FALSE}
per <- periodicity(extract,instr='Guitar',freq_range=c(0.1,0.6))
print(per$Figure)
```

## Summarise Periodicity 

Convenience functions to extract peak of periodicity curve. 

```{r}
psumm <- summarise_periodicity(per$Curve)
print(paste('Period is:', round(psumm$Per,3), 'seconds'))
```

Or convert to BPM

```{r}
print(paste('In BPM:',period_to_BPM(psumm$Per*2))) # knowing these are sub-beats
```

## Estimate Periodicity

Use **autocorrelation** this time...

```{r}
per2 <- periodicity(extract,instr='Guitar',method='acf',freq_range=c(0.1,0.6))
print(per2$Figure)
```


## Other Sources of Onset Data {.smaller}

### Real performances

  - [CompMusic datasets at UPF: Hindustani (North India), Carnatic (South India), Turkish-makam (Turkey), Arab-Andalusian (Maghreb), and Beijing Opera (China)](https://compmusic.upf.edu/datasets)
  - [Tap & Fiddle: Scandinavian Fiddle Tunes with Accompanying Foot-Tapping](https://zenodo.org/record/4308731)
  - [The MAESTRO (MIDI and Audio Edited for Synchronous Tracks and Organisation)](https://magenta.tensorflow.org/datasets/maestro)
    
    - 200 hours of real recordings with MIDI and audio (Piano competition)
  
  - [University of Rochester Multi-Modal Musical Performance](http://www2.ece.rochester.edu/projects/air/projects/URMP.html) 
  - [Carnatic Music Rhythm Dataset](http://compmusic.upf.edu/carnatic-varnam-dataset)
  

## Credits

::: callout-note
Thanks for **Adrian Poole** and **Simone Tarsitani** for preparing [IEMP Cuban](10.17605/osf.io/s437t) data included here.
:::


# References

Clayton, M., K. Jakubowski, T. Eerola, P. Keller, A. Camurri, G. Volpe, and P. Alborno. (2020). Interpersonal Entrainment in Music Performance: Theory, Method and Model. _Music Perception 38 (2)_, 136–94.

Clayton, M., S. Tarsitani, R. Jankowsky, L. Jure, L. Leante, R. Polak, A. Poole, et al. (2021). The Interpersonal Entrainment in Music Performance Data Collection. _Empirical Musicology Review 16 (1),_ 65–84. [http://dx.doi.org/10.18061/emr.v16i1.7555](http://dx.doi.org/10.18061/emr.v16i1.7555)

Rasch, Rudolf A. (1979). Synchronization in Performed Ensemble Music. _Acta Acustica United with Acustica 43 (2)_, 121–31.

Repp, Bruno H, and Yi-Huang Su. (2013). Sensorimotor Synchronization: A Review of Recent Research (2006–2012). _Psychonomic Bulletin & Review 20 (3)_, 403–52.

Wilkinson, Mark D, Michel Dumontier, IJsbrand Jan Aalbersberg, Gabrielle Appleton, Myles Axton, Arie Baak, Niklas Blomberg, et al. (2016). The FAIR Guiding Principles for Scientific Data Management and Stewardship. _Scientific Data 3 (1)_, 1–9.

Wing, Alan M, Satoshi Endo, Adrian Bradbury, & Dirk Vorberg. (2014). Optimal Feedback Correction in String Quartet Synchronization. _Journal of The Royal Society Interface 11 (93)_, 20131125.
