---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# onsetsync - Analysis and Visualisation of Synchronisation of Music Onset Data


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`onsetsync` is a R package for musical dynamics involving synchrony. There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property. 

## Installation

You can install the current version of `onsetsync` from Github by
entering the following commands into R:

```{r,eval=FALSE}
if (!require(devtools)) install.packages("devtools")
devtools::install_github("tuomaseerola/onsetsync")
```

## Usage

Note that `onsetsync` is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g. Librosa and Python, or Matlab, or Sonic Visualiser) using well-known algorithms. Here we take it as granted that we have extracted the onsets elsewhere and we have the onset times recorded into the csv files. 

```{r, results='hide', message=FALSE, warning=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(httr)
library(onsetsync)
```

```{r message=FALSE,warning=FALSE,eval=TRUE}
# Onsets
invisible(GET("https://osf.io/8a347/?action=download", 
              write_disk("CSS_Song2_Onsets_Selected.csv", overwrite = T)))
CSS_Song2_Onset<-read.csv("CSS_Song2_Onsets_Selected.csv",header=T,sep=",")
knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple",
             caption = 'Onset data structure (see text for details).')

# Annotations
invisible(GET("https://osf.io/4cdpr/?action=download", 
              write_disk("CSS_Song2_Metre.csv", overwrite = TRUE)))
CSS_Song2_Metre <- read.csv("CSS_Song2_Metre.csv", header=T, sep=",")
knitr::kable(head(CSS_Song2_Metre[1:4,]),format = ,"simple")
CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
                                 Label.SD,SD,Clave,Bass,Guitar,Tres) 
```

As the onsets and annotations are in different files, let's first combine the raw onset and annotation with `onsetsync`. Here we first add annotations about the cycles into the onset data and then we also add isochronous beat times to the data frame, since are useful reference points for future calculations.


```{r message=FALSE,warning=FALSE,eval=TRUE}
# Add annotations about the cycle to the data frame
CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
                            annotation = CSS_Song2_Metre$Cycle,
                            time = CSS_Song2_Metre$Time,
                            reference = 'Label.SD')
# Add isochronous beats to the data frame
CSS_Song2 <- add_isobeats(df = CSS_Song2, 
                          instr = 'CycleTime', 
                          beat = 'SD')

print(knitr::kable(head(CSS_Song2),format = "simple",digits = 2))
```


Before moving onto the analysis, let's summarise the onset structures in this piece.


```{r message=FALSE,warning=FALSE,eval=TRUE}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres'))
print(knitr::kable(tab1,digits = 1,
     caption = 'Descriptives for the onset time differences (ms)'))
```

s a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time. This gives a summary about when the instrument are playing relative to the beat division and the sections of the song. 

```{r synch2isochron,fig.width=9, fig.height=7.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual='Isochronous.SD.Time',
                     pcols=2)
print(fig1)
```


