---
title: "Analysis Example"
output: rmarkdown::html_vignette
description: >
  Analysis example showcasing the different types of analyses available.
vignette: >
  %\VignetteIndexEntry{Analysis Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=FALSE}
latex_format <- FALSE # set TRUE to compile the LaTeX paper
save_figures <- FALSE # set TRUE to compile the LaTeX paper
```

# Introduction

These analyses are the ones reported in paper titled "onsetsync: An R Package for Onset Synchrony Analysis" (submitted) by [Tuomas Eerola](https://www.durham.ac.uk/staff/tuomas-eerola/) and [Martin Clayton](https://www.durham.ac.uk/staff/martin-clayton/). The sections and figures correspond to the manuscript.

## Load libraries

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
devtools::install_github("tuomaseerola/onsetsync", force = TRUE)
library(onsetsync)
packageVersion("onsetsync")
```

## Read data

```{r,results='asis'}
data("CSS_IEMP", package = "onsetsync")
CSS_Song2 <- CSS_IEMP[[2]]
CSS_Song2 <- dplyr::select(CSS_Song2,Piece,Section,Clave,Bass,Guitar,Tres,SD,Cycle,Isochronous.SD.Time) # delete previous
print(knitr::kable(head(CSS_Song2),digits = 2,format = if(latex_format==TRUE){'latex'} else {'markdown'}))
```

## Summarise onsets

```{r, results='asis'}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres','Isochronous.SD.Time'))
print(knitr::kable(tab1, digits = 1,format = if(latex_format==TRUE){'latex'} else {'markdown'}))
```
### Figure 1

```{r, fig.width=4, fig.width=7.5}
print(plot_by_beat(df = CSS_Song2, 
                   instr = c('Bass','Clave','Guitar','Tres'), 
                   beat = 'SD', 
                   virtual='Isochronous.SD.Time',
                   pcols=2))
```

```{r, echo=FALSE}
if(save_figures==TRUE){
fig1<-plot_by_beat(df = CSS_Song2, 
                   instr = c('Bass','Clave','Guitar','Tres'), 
                   beat = 'SD', 
                   virtual='Isochronous.SD.Time',
                   pcols=2)
ggsave(filename = 'figure1.pdf',fig1,height = 4,width = 7.5)
}
```

# Analysis of Synchrony

Begin with a diagnostic plot.

### Figure 2
```{r, fig.height=5, fig.width=8}
print(plot_by_beat(df = CSS_Song2,
                     instr = c('Guitar','Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     griddeviations = TRUE))
```

```{r, echo=FALSE}
if(save_figures==TRUE){
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Guitar','Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     griddeviations = TRUE)
print(fig2)
ggsave(filename = 'figure2.pdf',fig2,height = 6,width = 10)
}
```

## Synchrony between the instruments

```{r}
d1 <- sync_sample_paired(CSS_Song2, instr1 = "Guitar", instr2 = "Tres", beat = "SD")
dplyr::summarise(data.frame(d1), N = n(), Mean.ms = mean(asynch*1000), Sd.ms = sd(asynch*1000))
```

### Figure 3

```{r, fig.width=4, fig.width=6}
inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
dn <- sync_execute_pairs(CSS_Song2,inst,beat = "SD")
print(plot_by_pair(dn))
```

```{r, echo=FALSE}
if(save_figures==TRUE){
  fig3 <- plot_by_pair(dn)  # plot
  print(fig3)  
  ggsave(filename = 'figure3.pdf',fig5,height = 4,width = 6)
}
```

### Plain summaries of asynchrony

Reporting of several measures of absolute asyncronies.

```{r}
d <- sync_sample_paired(CSS_Song2,'Clave','Bass',beat = 'SD')
print(summarise_sync(d))
```
Relative measure for two instruments vs other other instruments.

```{r}
x<-sync_sample_paired_relative(df = CSS_Song2,instr = 'Bass',instr_ref = c('Guitar','Bass','Tres','Clave'),beat = 'SD')
print(x$`Mean pairwise asynchrony`)

x <- sync_sample_paired_relative(df = CSS_Song2,instr = 'Clave',instr_ref = c('Guitar','Bass','Tres','Clave'),beat = 'SD')
print(x$`Mean pairwise asynchrony`)
```

### Asynchrony across beat subdivisions

```{r}
dn <- sync_execute_pairs(df = CSS_Song2, instruments = c("Guitar", "Tres"),beat = "SD")
print(summarise_sync_by_pair(dn, bybeat = TRUE))
```

## Synchrony across several performances

Using the built-in corpus of Cuban Salsa and Son preformances (`CSS_IEMP`).

```{r, fig.height = 3.5, fig.width = 6}
corpus <- onsetsync::CSS_IEMP
D <- sync_sample_paired(corpus,'Guitar','Tres',n=0,beat='SD')
D <- D$asynch
D$asynch_abs <- abs(D$asynch)*1000
print(paste('Mean asynchrony across all performances is', round(mean(D$asynch)*1000,1),'ms'))
```

### Figure 4

```{r, fig.height = 3.5, fig.width = 6}
print(plot_by_dataset(D,'asynch_abs','name', box = TRUE) + ylab('Absolute Asynchrony (ms)'))
```

```{r, echo=FALSE}
if(save_figures==TRUE){
  fig4 <- plot_by_dataset(D,'asynch_abs','name', box = TRUE)
  fig4 + ylab('Absolute Asynchrony (ms)')
  ggsave(filename = 'figure4.pdf',fig7,height = 3.5,width = 6)
}
```

## Synchrony with other variables

```{r,fig.height = 3,fig.width = 6.5}
CSS_Song2 <- CSS_Song2 %>%
  group_by(Cycle) %>%
  mutate(Tempo = 240/(max(Isochronous.SD.Time) - min(Isochronous.SD.Time)))

d2 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", beat = 'Tempo')
d3 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", beat = 'Isochronous.SD.Time')
tmp <- data.frame(asynch = d2$asynch*1000, Tempo = d2$beatL, Time = d3$beatL)
```

### Figure 5

```{r,fig.height = 3,fig.width = 6.5}
print(plot_by_var_time(df = tmp,
   var1 = 'Time',
   var2 = 'asynch',
   var3 = 'Tempo',
   ylabel = 'Asynchrony (ms)'))
```

```{r, echo=FALSE}
if(save_figures==TRUE){
  fig5 <- plot_by_var_time(df = tmp,
                         var1 = 'Time',
                         var2 = 'asynch',
                         var3 = 'Tempo',
                         ylabel = 'Asynchrony (ms)')

ggsave(filename = 'sync_across_time.pdf',fig5, height = 3,width = 6.5)
}
```

## Periodicity

### Figure 5

```{r,fig.height = 4.8, fig.width = 7.2}
extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
extract <- ungroup(extract) # drop grouping structure
fig10 <- gaussify_onsets(extract$Guitar, wlen = 0.2, plot = TRUE)

frq_range <- c(0.10,0.50)
per1 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
                    method = 'diff', title = 'Difference')
per2 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
                    method = 'acf',title = 'Autocorrelation')
per3 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
                    method = 'per',title = 'FFT')

plots <- align_plots(fig10$fig, align = 'v', axis = 'l')
bottom_row <- plot_grid(per1$Figure, per3$Figure, per2$Figure, 
                        labels = c('B', 'C', 'D'), label_size = 12,nrow =1)
fig5 <- plot_grid(plots[[1]], bottom_row, labels = c('A', ''), 
                  label_size = 12, ncol = 1)
print(fig5)
```

```{r, echo=FALSE}
if(save_figures==TRUE){
  ggsave(filename = 'figure5.pdf',G1,height = 4,width = 6)
}
```

### Report the peak periodicity

```{r}
PM <- summarise_periodicity(per3$Curve)
print(paste('Period of',round(PM$Per*1000),'ms'))

print(paste(period_to_BPM(PM$Per*4),'BPM'))
```

## Use of periodicity analysis

Take an excerpt from the example performance and apply periodicity analysis to four instruments.

### Figure 6

```{r, fig.height=4.8, fig.width=7.4}
excerpt <- dplyr::filter(CSS_Song2,Guitar>60 & Guitar < 60*2)
excerpt <- data.frame(excerpt)
p1 <- onsetsync::periodicity(excerpt,instr = "Guitar",
                             freq_range = c(0,1),sampling_rate = 100)
p2 <- onsetsync::periodicity(excerpt,instr = "Bass",
                             freq_range = c(0,1),sampling_rate = 100)
p3 <- onsetsync::periodicity(excerpt,instr = "Tres",
                             freq_range = c(0,1),sampling_rate = 100)
p4 <- onsetsync::periodicity(excerpt,instr = "Clave",
                             freq_range = c(0,1),sampling_rate = 100)
p1$Curve$Instrument <- "Guitar"
p2$Curve$Instrument <- "Bass"
p3$Curve$Instrument <- "Tres"
p4$Curve$Instrument <- "Clave"
df<-rbind(p1$Curve,p2$Curve,p3$Curve,p4$Curve)
df$Instrument<-factor(df$Instrument)

fig6 <- ggplot(df,aes(x=Time,y=Ampl))+
  geom_line(color='blue')+
  facet_wrap(.~Instrument,scales = "free",nrow = 2)+ # ,
  theme_linedraw()+
  scale_y_continuous(expand = c(0.01,0.01))+
  scale_x_continuous(breaks = seq(0,1,by=0.217/2),
                     labels = scales::number_format(accuracy = 0.01),expand = c(0.01,0.01))+
  xlab('Period (s)')+
  ylab('Amplitude')
print(fig6)
```

```{r, echo=FALSE}
if(save_figures==TRUE){
  ggsave(filename='figure6.pdf', g1, height = 5, width = 8)
}
```

