---
title: "onsetsynch"
author: "Tuomas Eerola"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{synchrony}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is a simple package meant to facilitate the analysis of onset data. Made in the context of Interpersonal Entrainment Music Performance (IEMP).

## Example

We will first read in data which consists of extracted onsets (done with MIR toolbox for Matlab) and hand annotated labelling of beats and structures. This is a Cuban song, recorded by Adrian Poole (more details here).

### Read data

```{r}
asere <- read.csv(url('https://raw.githubusercontent.com/tuomaseerola/onsetsynch/master/data/Asere_OU_2.csv'))
head(asere)
```

There is quite a bit of extra data here, which we won't need. I also want to handle the data in `tidyverse` fashion and give few diagnostics about the data.

```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
```

```{r}

asere <- as_tibble(asere)
asere <- select(asere,SD,Section,Tempo,SD,Virtual.SD,Clave,Bass,Guitar,Tres,Bongo,Bell) # keep only key variables
print(dim(asere))
round(colSums(!is.na(asere[,which(colnames(asere)=="Clave"):which(colnames(asere)=="Bell")]),na.rm = T)/nrow(asere),2) # percentage of available data
```

### Load `onsetsynch` library

OK. Load up the library and visualise the relative asynchrony to equal division for each instrument.

```{r fig.width=7.5, fig.height=7.5}
devtools::install_github("tuomaseerola/onsetsynch") # install the developmental package synchrony
library(onsetsynch)
p <- onsetsynch_by_beat_plot(asere,c('Bass','Clave','Guitar','Tres','Bongo','Bell'),'SD','Virtual.SD',pcols=2)
print(p)
```

### Calculate the synchrony between the instruments

How well are the pairs of instruments synced? Since the instrument play different amounts of onset, and these are bound to at different beats, the mutual amount of onsets for each pair varies. In order to keep the mean and standard deviations comparable, we will randomnly sample the joint onsets for both instruments. 

```{r}
set.seed(1201) # set random seed before doing any sampling in order replicate results
N <- 100 # Let's select 100 onsets
d1 <- onsetsynch_sample_paired(asere,'Clave','Bass',N,1,'SD',TRUE) # Determine the instrument pair
print(paste('Asynchrony mean=',round(mean(d1$asynch*1000),4),'ms & sd. dev=',round(sd(d1$asynch*1000),3)))
```
The first example might still be in accurate since we now know that there are at least 241 shared onset times between the Clave and the Bass. Let's redo the random sampling 10 times so we get more observations whilst still always sampling 100 joint onsets.
```{r}
d10 <- onsetsynch_sample_paired(asere,'Clave','Bass',N,10,'SD',TRUE) # New parameter 10 for 10 times
print(paste('Asynchrony mean=',round(mean(d10$asynch*1000),4),'ms & sd. dev=',round(sd(d10$asynch*1000),3)))
```

Carry this out for all possible pairings of the instruments and visualise the results.

```{r fig.width=7.5, fig.height=7.5}
inst<-c('Clave','Bass','Guitar','Tres','Bongo','Bell') # Define instrument for exhaustive pairwise comparison
dn <- onsetsynch_execute_pairs(asere,inst,N,10,'SD')
dim(dn$asynch)

p2 <- onsetsynch_by_pair_plot(dn)  # plot
print(p2)  

# Compare only three interesting instruments in terms of the synchrony across the beats
inst<-c('Bass','Guitar','Tres') # Define instrument for exhaustive pairwise comparison
dn <- onsetsynch_execute_pairs(asere,inst,N,10,'SD')

p3 <- onsetsynch_by_pair_plot(dn,bybeat=TRUE) # plot by beats
print(p3)  
```
Calculate simple t-statistics for the instrument pairs.
```{r}
T <- data.frame(onsetsynch_by_pair_stats(dn))
print(T)
```
### Other visualisations

Visualise asynchronies against another variable, for instance, tempo.

```{r fig.width=5.5, fig.height=5.5}
d1 <- onsetsynch_sample_paired(asere,'Clave','Bass',200,1,'Tempo')
g1 <- onsetsynch_by_X_plot(d1,meta = 'Clave-Bass',xlab='Tempo (BPM)')
print(g1)
```


