---
title: "Synthesise onsets"
output: rmarkdown::html_vignette
description: >
  Create audio files from the onset data.
vignette: >
  %\VignetteIndexEntry{Synthesise onsets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To create an audio file of your onset data, you can call `synthesise_onsets` function.

## Example data

Load a Cuban salsa example. Let's take a 10 second example. To eliminate the first 5, where the instruments are not playing, let's redefine the time to start from the first annotated cycle.

```{r load, warning=FALSE,message=FALSE}
library(onsetsync)
data <- CSS_IEMP[[2]]                                                          # example
data <- dplyr::filter(data,Isochronous.SD.Time > 5 & Isochronous.SD.Time < 15) # 10 s
data$Guitar <- data$Guitar - min(data$Isochronous.SD.Time,na.rm = TRUE)        # trim time     
data$Bass <- data$Bass - min(data$Isochronous.SD.Time,na.rm = TRUE)            # trim time
```

## Synthesise Guitar and 

Let's take two instruments and create their onsets as high-pitched sound (Guitar) and short noise burst (Bass).

```{r createwave, echo=TRUE}
inst <- c('Guitar','Bass')
fs <- 22050
w <- synthesise_onsets(data=data,instruments = inst, sr = fs, type =c('synth','noise'))
```

Plot the signal. This is downsampled for plotting in `ggplot2`.

```{r, echo=TRUE,fig.width=6.5,fig.height=2.5}
scaling <- 20
ws <- signal::resample(w,fs,fs*scaling)
tmp <- data.frame(Amplitude=ws,Time=seq(0,length(ws)-1))
tmp$Time <- tmp$Time/fs*scaling
ggplot2::ggplot(tmp,ggplot2::aes(x=Time,y=Amplitude))+
  ggplot2::geom_line(color='blue')+
  ggplot2::theme_linedraw()
```

Save as wave file using `seewave` library.

```{r savefile, echo=TRUE, eval=FALSE}
seewave::savewav(wave = w, f = fs, channel = 1, filename = 'example.wav')
```

Further options relate to sound generators and frequency.
